Turbo fast no compromise gravitational wave parameter estimation

### 2022-11-21 src/tex/ms.tex
update

\definecolor{rb4}{HTML}{27408B}
\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}

@@

% Author list
\author{@kazewong}

% Abstract with filler text
\begin{abstract}
We present a light-weighted, flexible and high performance framework to estimate
gravitational wave event parameters. By combining heterodyned likelihood,
automatically differentiable and accelerators compatible waveforms, normali zing
flow enhanced gradient-based MCMC, we achieve parameter estimation for real
event like GW150914 and GW170817 within a minute of sampling time with little to
no compromise. Our framework does not require pre-training or explicit
reparameterization, and can be generalized to handle higher dimensional
problems. We also discuss as near-real time parameter estimation have been shown
to be possible by multiple groups, what are the trade-offs and future
developments the community should be aware of. Our code for generating the
manuscript and running the analysis is publically available on github.
\end{abstract}

@@

% Brief description of PE and old generation of code
Parameter estimation (PE) is one of the most commonly perform data analysis
tasks in gravitational wave (GW) data analysis. The central question of PE is to
infer the parameters of the GW source given the strain data. The PE results in
the most recent catalog of GW events are powered by a number of community
developed PE codes, including \texttt{lalinference}, \texttt{pycbc}, and
\texttt{bilby} \kw{cite}. These packages have been tested by a number of groups
and well regarded as the standard tools.

@@

@@

In this section we give a self-contained summary the gravitational-wave
parameter estimation process.

@@

\begin{align}
    p(\theta| d) = \frac{\mathcal{L}(d|\theta)\pi(\theta)}{p(d)}
\end{align}

\begin{align}
    \log{\mathcal{L}(d|\theta)} = -\left<d-h|d-h\right>/2
\end{align}

\begin{align}
    \left<a|b\right> = 4 Re\int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)} df
\end{align}

>>>>

\definecolor{rb4}{HTML}{27408B}
\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
\newcommand{\flatiron}{\affiliation{Center for Computational Astrophysics,
Flatiron Institute, New York, NY 10010, USA}}
\newcommand{\jhu}{\affiliation{Department of Physics and Astronomy, Johns Hopkins University,
3400 N. Charles Street, Baltimore, Maryland 21218, USA}}

@@

% Author list
\author{Kaze W. K. Wong}
\email{kwong@flatironinstitute.org}
\flatiron

\author{Thomas Edward}
\jhu

\author{Maximiliano Isi}
\flatiron

% Abstract with filler text
\begin{abstract}
We present a light-weighted, flexible and high performance framework to estimate
gravitational wave event parameters. By combining heterodyned likelihood,
automatically differentiable and accelerators compatible waveforms, normalizing
flow enhanced gradient-based MCMC, we achieve parameter estimation for real
event like GW150914 and GW170817 within a minute of sampling time with little to
no compromise. Our framework does not require pre-training or explicit
reparameterization, and can be generalized to handle higher dimensional
problems. We also discuss as near-real time parameter estimation have been shown
to be possible by multiple groups, what are the trade-offs and future
developments the community should be aware of. Our code for generating the
manuscript and running the analysis is publically available on github.
\end{abstract}

@@

% Brief description of PE and old generation of code
Parameter estimation (PE) is one of the most commonly perform data analysis
tasks in gravitational wave (GW) data analysis. The central question of PE is to
infer the parameters of the GW source given the strain data. The PE results in
the most recent catalog of GW events are powered by a number of community
developed PE codes, including \texttt{lalinference}, \texttt{pycbc}, and
\texttt{bilby} \cite{Romero-Shaw:2020owr} \kw{cite}. These packages have been tested by a number of groups
and well regarded as the standard tools.

@@

\kw{Dingo can't do BNS yet}

@@

In this section we give a self-contained summary the gravitational-wave
parameter estimation (PE) process.

@@

The main objective in PE can be summarized as: given data in the form of a time
series of strain $\mathbf{d}$, find the distribution of the source parameters
$\mathbf{\theta}$ that is compatible with the data, i.e.
$p(\mathbf{\theta}|\mathbf{d})$. In order to compute this object, we can rewrite
it using Bayes' theorem,
\begin{align}
    p(\theta| d) = \frac{\mathcal{L}(d|\theta)\pi(\theta)}{p(d)},
\end{align}

where $\mathcal{L}(d|\theta)$ is the likelihood function, $\pi(\theta)$ is the
prior distribution, and $p(d)$ is the evidence. Since the evidence is a
normalization constant that does not depend on the source parameters, it is
often omitted if we are only interested in the posterior distribution. The prior
distribution is often chosen to be some simple distributions, such as uniformly
distributed in the component masses or a Gaussian distribution in the spins.
The most computationally intensive part of the PE is the evaluating the likelihood function.
Assuming the noise follows from a stationary Gaussian process, the log likelihood in the case of GW is given by

\begin{align}
    \log{\mathcal{L}(d|\theta)} = -\left<d-h(\mathbf{\theta})|d-h(\mathbf{\theta})\right>/2,
\end{align}

where $d$ is the data, $h(\mathbf{\theta})$ is the strain predicted by a waveform model with a specific set of source parameters.
It is computationally cheaper to compute the likelihood in the frequency domain, which can be defined as 
\begin{align}
    \left<a|b\right> = 4 Re\int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)} df,
\end{align}
where $\mathcal{S}_n(f)$ is the noise power spectral density (PSD).
